Quantum Mechanics Assignment Nine

### 2016-03-06 quantum/latex/a9.tex
start final quantum assignment.

\title{Quantum Mechanics Assignment Eight}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem 1: Two Spin-1 Particle states}
Here we use a Clebsch-Gordan table and our knowledge of the P interchange operator
to list the spin states a system of two spin-1 particles could take on, and whether
they are symmetric or antisymmetric with respect to particle interchange.\\

Two spin particles combine to have a total angular momentum of 0, 1 or 2. We use a CG
table to list the states, in the format $\ket{J,M_J} = \ket{M_{S1}, M_{S2}}$:\\

\begin{align*}
  \mbox{Symmetric}\hspace{2cm} &\ket{2,2} = \ket{1,1} = \ket{1,1}\\
  \mbox{Symmetric}\hspace{2cm}&\ket{2,1} = \frac{1}{\sqrt{2}}\left(\ket{1,0}
  + \ket{0,1}\right)
  = \frac{1}{\sqrt{2}}\left(\ket{0,1} + \ket{1,0}\right)\\
  \mbox{Symmetric}\hspace{2cm} &\ket{2,0} =
  \frac{1}{\sqrt{2}}\left(\ket{1,-1} + \ket{-1,1}\right) + \sqrt{\frac{2}{3}}\ket{0,0}
  = \frac{1}{\sqrt{2}}\left(\ket{-1,1} + \ket{1,-1}\right)
  + \sqrt{\frac{2}{3}}\ket{0,0}\\
  \mbox{Symmetric}\hspace{2cm} &\ket{2,-1} =
  \frac{1}{\sqrt{2}}\left(\ket{-1,0} + \ket{0,-1}\right)
  = \frac{1}{\sqrt{2}}\left(\ket{0,-1} + \ket{-1,0}\right)\\
  \mbox{Symmetric}\hspace{2cm} &\ket{2,-2} = \ket{-1,-1} = \ket{-1,-1}\\
\end{align*}

\begin{align*}
  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,1}
  = \frac{1}{\sqrt{2}}\left(\ket{1,0} - \ket{0,1}\right)
  = -\frac{1}{\sqrt{2}}\left(\ket{0,1} - \ket{1,0}\right)\\
  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,0}
  = \frac{1}{\sqrt{2}}\left(\ket{1,-1} - \ket{-1,1}\right)
  = -\frac{1}{\sqrt{2}}\left(\ket{-1,1} - \ket{1,-1}\right)\\
  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,-1}
  = \frac{1}{\sqrt{2}}\left(\ket{-1,0} - \ket{0,-1}\right)
  = -\frac{1}{\sqrt{2}}\left(\ket{0,-1} - \ket{-1,0}\right)\\
\end{align*}

\begin{align*}
  \mbox{Symmetric}\hspace{2cm} &\ket{0,0} = \ket{-1,-1} =
  \frac{1}{\sqrt{3}}\left(\ket{1,-1} - \ket{0,0} \ket{-1,1}\right)
  = \frac{1}{\sqrt{3}}\left(\ket{-1,1} - \ket{0,0} \ket{1,-1}\right)\\
\end{align*}

This is a demonstration of how the coupled basis states for 2x spin-1 particles
obey the Symmetrization Postulate. This is an advantage that the uncoupled basis
does not share: $\ket{1,-1}$ does not obey the SP, since the first particle is always
in the $M_S = 1$ state, and so is effectively ``labelled''. This is disallowed
by QM.\\

An interesting result we can glean from this is that two fermions with
spin-1 (eg protons) will always go to the F=1 state when they have a symmetric
wave function (ie they are in the ground state), since all wavefunctions must be
antisymmetric for fermions. This is a weird case that is not intuitive - why is
there such an arbitrary set of rules on what the spin can and cannot be?\\

\section{Problem 2: Neutron Star}
Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.
They are incredibly massive, and without the Pauli Exclusion Principle would overcome the
inter-particle repulsions and collapse into a black hole. However, because of the PEP, no two
particles in a system can take on the same quantum state. I don't know the quantum numbers of
a neutron in a system like this, but I do know that no two neutrons in the system are allowed to
have the same spacial wave function, because this would violate the PEP. This requirement for each
particle to have a different spacial wave function translates to the particles needing to be in
different spaces, and hence not collapsing into the center and forming a black hole. Thus there
is a ``quantum force'' which opposes gravity and keeps the particles stable.\\

This is a really weird thing to think about. It is odd that the entire star is treated as a single
system, meaning each neutron must have different quantum numbers, since this translates to an
immense number of numbers. It is odd that the PEP manifests itself as a force which opposes gravity.
\\

Another rival explanation is that, due to the Heisenberg Uncertainty Principle, the position of
each particle is well known due to tight packing, so the momentum uncertainty must be very large.
This translates to the particles having a large range of momenta. This momenta could possibly be
the means by which the particles oppose the gravitational force, since the ``quantum force''
explanation seems weird to me.\\

\section{Problem 13.7: Noninteracting particles in a HO potential}
Here we find the interparticle spacing for three types of systems: a distinguishable,
fermonic and bosonic pair of particles in a QHO potential where one particle is in state $n$
and the other in state $k$. We construct our states as follows, noting that fermions have AS
descriptions, bosons have S descriptions and distinguishable particles have neither:

\begin{align*}
  \mbox{Distinguishable}&\hspace{1cm} \ket{\psi} = \ket{n}_1\ket{k}_2\\
  \mbox{Fermions}&\hspace{1cm} \ket{\psi}
  = \frac{1}{\sqrt{2}}\left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
  \mbox{Bosons}&\hspace{1cm} \ket{\psi}
  = \frac{1}{\sqrt{2}}\left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
\end{align*}

\textbf{a.) Distinguishable}

\begin{align*}
  \braket{(x_1-x_2)^2} &= \bra{n}_1\bra{k}_2(x_1-x_2)^2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1\bra{k}_2(x_1^2 + x_2^2 - 2x_1x_2)\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1\bra{k}_2x_1^2\ket{n}_1\ket{k}_2 +
  \bra{n}_1\bra{k}_2x_2^2\ket{n}_1\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1x_1^2\ket{n}_1\bra{k}_2\ket{k}_2 +
  \bra{n}_1\ket{n}_1\bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\bra{n}_1
  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{n}_1
  + \frac{\hbar}{2m\omega}\bra{k}_2
  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{k}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\bra{n}_1\left(aa^\dagger + a^\dagger a\right)\ket{n}_1
  + \frac{\hbar}{2m\omega}\bra{k}_2\left(aa^\dagger + a^\dagger a\right)\ket{k}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
  \left(a_1+a_1^\dagger\right)\left(a_2+a_2^\dagger\right)\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\bra{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)\\
  &- \frac{2\hbar}{2m\omega}
  \left(\bra{n}_1a_1\ket{n}_1\ket{k}_2a_2\ket{k}_2
  + \bra{n}_1a_1^\dagger\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
  + \bra{n}_1a_1\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
  + \bra{k}_2a_2\ket{k}_2\bra{n}_1a_1^\dagger\ket{n}_1\right)\\
  &= \frac{\hbar}{m\omega}(n+k+1)\\
\end{align*}

\textbf{b.) Boson}

\begin{align*}
  \braket{(x_1-x_2)^2} &=
  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 + \bra{k}_1\bra{n}_2\right)
  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
  \left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
  &= \frac{1}{2}\left(
  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
  + \bra{k}_1x_1^2\ket{k}_1 + \bra{n}_2x_2^2\ket{n}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
  &= \frac{1}{2}\left(4n + 4k + 4\right)
  - \bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - \bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\\
  &= \frac{\hbar}{4m\omega}\left(4n + 4k + 4\right)\\
  &= \frac{\hbar}{m\omega}\left(n + k + 1\right)\\
\end{align*}

\textbf{c.) Fermion}

\begin{align*}
  \braket{(x_1-x_2)^2} &=
  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 - \bra{k}_1\bra{n}_2\right)
  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
  \left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
  &= \frac{1}{2}\left(
  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
  - \bra{k}_1x_1^2\ket{k}_1 - \bra{n}_2x_2^2\ket{n}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 + 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
  &= \frac{\hbar}{4m\omega}\left(2n+1 + 2k+1 - 2n-1 - 2k-1 - 0\right)\\
  &= 0\\
\end{align*}

Note: These solutions assume that n and k are not two levels separate of each other.\\

This is a really interesting problem that demonstrates how systems that are identical save
for their symmetry rules can behave totally differently. The boson system has twice the expectation
value of the distinguishable particle system, and the fermion system has zero expectation value
for particle distance. This is a very unintuitive result, since it predicts that on average the
two particles are on top of each other.\\

\section{Problem 4: Three-particle system}
In this problem we practice writing the wave functions for three-particle systems of particles
with different symmetry rules. Here A, B and C represent spacial states and $S_A, S_B$ and $S_C$ are
spin states. Particles 1, 2 and 3 are the three particles in the system.\\

\textbf{Distinguishable Particles}

The distinguishable particle case is simple. Because we are allowed to label particles, we say that
each particle has its own spacial and spin states A, B and C. There is no need for us to make our
wave function have equal probabilities of measuring 1, 2 and 3 in any state, we are allowed to say
particle 1 is in state A:\\

\begin{align*}
  \Psi_D &= \left(\Psi_1^A\Psi_2^B\Psi_3^C\right)\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3\\
\end{align*}

\textbf{Bosons and Fermions}

The calculation for the boson and fermion states is more difficult. Our wave functions must reflect
that measuring a particle in state A cannot tell us whether this particle is particle 1, 2 or 3. To
achieve this we sum all the possible combinations of states to create a completely ambiguous
superposition state. This is the symmetric state.\\

\begin{align*}
  \Psi^{S} &=
  \Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B\\
  &+ \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A\\
  &+ \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\\
  \ket{S}^{S} &=
  \ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3\\
  &+ \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3\\
  &+ \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\\
\end{align*}

We then use a Slater determinant to construct the antisymmetric spacial and spin states:\\

\begin{align*}
  \Psi^{A} &= \det
  \begin{vmatrix}
    \Psi_1^A & \Psi_2^A & \Psi_3^A\\
    \Psi_1^B & \Psi_2^B & \Psi_3^B\\
    \Psi_1^C & \Psi_2^C & \Psi_3^C\\
  \end{vmatrix}\\
  &= \Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B\\
  &- \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B\\
  &+ \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\\
  \ket{S}^A &= \det
  \begin{vmatrix}
    \ket{S_A}_1^A & \ket{S_A}_2 & \ket{S_A}_3\\
    \ket{S_B}_1^B & \ket{S_B}_2 & \ket{S_B}_3\\
    \ket{S_C}_1^C & \ket{S_C}_2 & \ket{S_C}_3\\
  \end{vmatrix}\\
  &= \ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3\\
  &- \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3\\
  &+ \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\\
\end{align*}

To create our boson wave functions, we take any combination of individual wave functions that result
in an overall symmetric wave function:\\

\begin{align*}
  \Psi_B^{SS} &= \Psi_B^S\ket{S}^S\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B
  + \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A
  + \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3
  + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)
\end{align*}

\begin{align*}
  \Psi_B^{AA} &= \Psi_B^A\ket{S}^A\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B
  - \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B
  + \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  - \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
\end{align*}

To make our fermion states, we take all the possible antisymmetric overall wavefunctions:\\

\begin{align*}
  \Psi_F^{AS} &= \Psi^A\ket{S}^S\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B
  - \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B
  + \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3
  + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
\end{align*}

\begin{align*}
  \Psi_F^{SA} &= \Psi^S\ket{S}^A\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B
  + \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A
  + \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  - \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
\end{align*}

\textbf{Many-electron system}
To describe a many-electron system, we use the following sets:\\

\begin{align*}
  A_{spacial} &= [\Psi^A, \Psi^B, ...]\\
  A_{spin} &= [\ket{S_A}, \ket{S_B}, ...]\\
\end{align*}

We also use function P(q), which returns all possible combinations of the elements of set q. Our
many-electron system is then represented by:\\

\begin{align*}
  \Psi_F^{AS} &= \Psi^A\ket{S}^S\\
  &= \left(\sum_p^{P(A_{spacial})}\prod_{i=0}^N p(i)_i\right)\det
  \begin{vmatrix}
    \ket{S_A}_1 & \ket{S_A}_2 & ... & \ket{S_A}_N\\
    \ket{S_B}_1 & \ket{S_B}_2 & ... & \ket{S_B}_N\\
    \vdots & \vdots & \vdots & \vdots\\
    \ket{S_N}_1 & \ket{S_N}_2 & ... & \ket{S_N}_N\\
  \end{vmatrix}
\end{align*}

\begin{align*}
  \Psi_F^{SA} &= \Psi^S\ket{S}^A\\
  &= \det
  \begin{vmatrix}
    \Psi^A_1 & \Psi^A_2 & ... & \Psi^A_N\\
    \Psi^B_1 & \Psi^B_2 & ... & \Psi^B_N\\
    \vdots & \vdots & \vdots & \vdots\\
    \Psi^N_1 & \Psi^N_2 & ... & \Psi^N_N\\
  \end{vmatrix}
  \left(\sum_p^{P(A_{spin})}\prod_{i=0}^N p(i)_i\right)
\end{align*}

This is a cool problem which made me think about how to generalize wave functions given the
Symmetrization Principle. Using Slater determinants was really helpful here, I'm not sure I could
have done the problem without them. My solution for representing many-electron systems is a bit
impractical, I'm interested in if and how such systems actually are represented.\\

\section{Problem 5: Zeeman Effect}
When doing Zeeman Effect calculations and deciding whether the Zeeman corrections should go in the
first-order Hamiltonian correction or be put in the unperturbed Hamiltonian, a ``strong'' B-field
is one that is too big to be a first-order correction. For the n=2 state of Hydrogen, the
first-order fine structure corrections have a size of:

\begin{equation*}
  E_{fs}^{(1)}
  = -\frac{1}{2}\alpha^4mc^2\frac{1}{2^3}\left(\frac{1}{j+\frac{1}{2}}-\frac{3}{8}\right)\\
  \approx -\frac{1}{16}\alpha^4mc^2\\
\end{equation*}

We say that Zeeman corrections of this size and larger should be considered due to ``strong''
magnetic fields. The energy perturbation due to the B-field is given by:\\

\begin{equation*}
  \Delta E_{Z} = \frac{\mu_B B}{\hbar}\left(g_\ell L_z + g_e S_z\right)
  \approx \frac{\mu_B B}{\hbar}\\
\end{equation*}

We equate these expressions to find the smallest ``strong'' B-field:\\

\begin{align*}
  B &\approx \frac{-\hbar}{16mu_B}\alpha^4mc^2
  = \frac{-6.5*10^{-16}}{16*6*10^{-5}}*7^4*10^{-12}*5*10^5\\
  \approx 8*10^{-26}T\\
\end{align*}

We now find the B-field which perturbes the H atom enough to equate the energies of the 2p and 3d
states:\\

\begin{align*}
  E_{2p} + \Delta E_{Zeeman, 2p} + E_{fs}^{(1)} &= E_{3d} + \Delta E_{Zeeman, 3d} + E_{fs}^{(1)}\\
  \frac{-13.6eV}{4} + \mu_BB(m_\ell+2m_s)
  + \frac{1}{2}\alpha^4mc^2\left(\frac{3}{4n^3}
  -\frac{\ell(\ell+1)-m_\ell m_s}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\right)
  &=\\ \frac{-13.6eV}{8} + \mu_BB(m_\ell+2m_s) + \frac{1}{2}\alpha^4mc^2\left(\frac{3}{4n^3}
  -\frac{\ell(\ell+1)-m_\ell m_s}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\right)\\
  \frac{-13.6eV}{4} + 2\mu_BB
  + \alpha^4mc^2\left(\frac{3}{64}
  -\frac{2-\frac{1}{2}}{24}\right)
  &= \frac{-13.6eV}{8} + -3\mu_BB + \frac{1}{2}\alpha^4mc^2\left(\frac{3}{108}
  -\frac{6+1}{405}\right)\\
  \frac{13.6eV}{4} + 5\mu_BB
  &= \alpha^4mc^2
  \left(\frac{3}{216} - \frac{7}{405} - \frac{3}{64} + \frac{3}{48}\right)\\
  B &= \frac{\alpha^4mc^2
    \left(\frac{3}{216} - \frac{7}{405} - \frac{3}{64} + \frac{3}{48}\right) - \frac{13.6eV}{4}}{5\mu_B}\\
  &= \frac{\frac{317\alpha^4mc^2}{25920} - \frac{13.6eV}{4}}{5\mu_B}\\
  &\approx \frac{0.68eV}{5.7*10^{-5}eV/T} = 10^4 T\\
\end{align*}

This is a huge magnetic field! I'm not entirely sure of it, but perhaps a very strong field is
necessary to split the orbitals to the level that they agree. It's important to understand when
the Zeeman effect can be treated as a perturbation and when it has to be actually included in
the original Hamiltonian. I said that whenever the Zeeman term is of a similar magnitude to the
fine structure correction, it is too big to include as a perturbation.\\

>>>>

\title{Quantum Mechanics Assignment Nine}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One: Gaussian Perturbation of Infinite Square Well}
Here we deal with an object in an infinite square well subject to a perturbation $V_0x^2e^{-t\tau}$.
The particle begins at $t=0$ in the ground state $E_1$. We wish to find the probability of a
transition to $E_2$ as a function of time $t$. To do this, we first calculate the perturbed
state coefficient $c_2(t)$:

\begin{align*}
  c_2(t) &= \frac{1}{i\hbar}\int_0^t\bra{2}H'\ket{1}e^{i(E_2-E_1)t'/\hbar}dt'\\
  &= \frac{2}{i\hbar L}\int_0^t
  \left(\int_0^L\sin\left(\frac{2\pi x}{L}\right)
  V_0x^2e^{-t/\tau}\sin\left(\frac{\pi x}{L}\right)dx\right)
  e^{\frac{i\hbar\pi^2(2^2-1^2)}{2mL^2}t}dt'\\
  &= \frac{2V_0}{i\hbar L}\int_0^t
  \left(\int_0^L\sin\left(\frac{2\pi x}{L}\right)
  x^2\sin\left(\frac{\pi x}{L}\right)dx\right)
  e^{-t/\tau}e^{\frac{i3\hbar\pi^2}{2mL^2}t}dt'\\
  &= \frac{2V_0}{i\hbar L}\int_0^t\frac{-8L^3}{\pi^2}e^{-t/\tau}e^{\frac{i3\hbar\pi^2}{2mL^2}t}dt'\\
  &= \frac{-16V_0L^2}{i\hbar\pi^2}
  \int_0^te^{\left(\frac{i3\hbar\pi^2}{2mL^2} - \frac{1}{\tau}\right)t}dt'\\
  &= i\frac{16V_0L^2}{\hbar\pi^2}\left(\frac{i3\hbar\pi^2}{2mL^2} - \frac{1}{\tau}\right)^{-1}
  e^{\frac{i3\hbar\pi^2}{2mL^2}t}e^{-\frac{t}{\tau}}\\
\end{align*}

We then find the probability of measurement:

\begin{align*}
  P_{1\rightarrow2} &= |c_2(t)|^2\\
  &= \left(\frac{16V_0L^2}{\hbar\pi^2}\right)^2
  \left(\frac{i3\hbar\pi^2}{2mL^2} - \frac{1}{\tau}\right)^{-2}
  e^{i\frac{6\hbar\pi^2}{2mL^2}t}e^{-2\frac{t}{\tau}}\\
\end{align*}

Because

\begin{align*}
  \lim_{t\rightarrow\infty} P_{1\rightarrow2}(t) = 0\\
\end{align*}

we can say that it becomes less and less likely for the system to be measured with energy
$E_2$ as time goes on.\\

This was a cool introduction to applying time-dependent perturbation theory equations to solving
problems. The answer is intuitive. The perturbation decays exponentially, so at large times it
should behave exactly like the unperturbed infinite square well and stay in the ground state with
100\% probability.\\
